Concurrent Programming with CXL

### 2020-06-07 book/paper.tex
Starting on monitors

While correct, it is not considered a good solution.
The solution is an example of what is called \emph{busy-waiting}:
processes spin in a loop until some desirable application-level condition is met.

@@

Tony Hoare, who came up with the concept of split binary semaphores, devised
an abstraction of the concept in a programming language paradigm called
\emph{monitors}.

>>>>

While correct, it is not considered a good solution.
The solution is an example of what is called \emph{busy-waiting}
(aka spin-waiting):
processes spin in a loop until some desirable application-level condition is met.

@@

Tony Hoare, who came up with the concept of split binary semaphores, devised
an abstraction of the concept in a programming language paradigm called
\emph{monitors}.
A monitor is a special version of an object-oriented \emph{class}, comprising
a set of variables and methods that operate on those variabled.
There is a split binary semaphore associated with each such class.
The \texttt{mutex} is hidden: it is automatically acquired when invoking a
method and released upon exit.
The other semaphores are called \emph{condition variables}.
There are two operations on condition variables: \texttt{wait} and \texttt{signal}.
If \texttt{c} is a pointer to a condition variable, then the semantics of the
operations, in CXL, are as follows:

\begin{code} 
\begin{verbatim} 
def wait(c):
    call V &(mutex);
    call P c;
;
def signal(c):
    call V c;
    call P &(mutex);
;
\end{verbatim} 
\end{code}

\texttt{wait} releases the monitor mutex and blocks while trying
to procure the condition variable (which is a semaphore that is 0 when
\texttt{wait} is invoked).
\texttt{signal} vacates the condition variable, passing the baton to
a process trying to procure it, and then procures the monitor mutex.
The use of \texttt{wait} and \texttt{signal} ensures that \texttt{P} and
\texttt{V} operations alternate as prescribed by the SBS technique.
Our implementations of reader/writer locks can be easily changed to
use \texttt{wait} and \texttt{signal} instead of using \texttt{P}
\texttt{V}.

In the late 70s, Xerox PARC developed a new programming language called
Mesa.  Mesa introduced various important concepts to programming languages,
including software exceptions and incremental compilation.  Mesa also
incorporated a version of monitors.
However, there are some subtle but important differences with Hoare-style
monitors that make Mesa-style monitors quite unlike split binary semaphores.
